Fitting Synchrotron Models to Gamma-Ray Burst Data from INTEGRAL/SPI

### 2021-12-17 src/ms.tex
rename bib file
change legend

\begin{figure}
    \begin{centering}
        \includegraphics{figures/lightcurve.pdf}
        \caption{Lightcurve for SPI detector 0 for GRB120711A. The transient source is clearly visible as well as the constant background for the time when the transient is not active.}
        \label{fig:lightcurve}
    \end{centering}
\end{figure}

@@

\begin{equation}
	L(\theta, b_{i}|D_{i}, B_{i},t_{d},t_{b}) = \frac{(t_{d}(m_{i}(\theta)+ b_{i}))^{D_{i}}}{D_{i}!}\cdot e^{-t_{d}(m_{i}(\theta)+b_{i})} \frac{(t_{b} b_{i})^{B_{i}}}{B_{i}!}\cdot e^{-t_{b} b_{i}}
  \label{eq:likelihood_m_and_b}
\end{equation}

Now we can marginalize over the parameters $b_{i}$, that we are not interested in, which leaves us with equation \ref{eq:likelihood_marg}.

\begin{equation}
	L(\theta|D_{i}, B_{i},t_{d},t_{b}) = \int_{0}^{\infty}\textrm{db}_{i}\frac{(t_{d}(m_{i}(\theta)+ b_{i}))^{D_{i}}}{D_{i}!}\cdot e^{-t_{d}(m_{i}(\theta)+b_{i})} \frac{(t_{b} b_{i})^{B_{i}}}{B_{i}!}\cdot e^{-t_{b}b_{i}}
  \label{eq:likelihood_marg}
\end{equation}

Because there is no easy analytical solution for this integral, one often uses a so-called profiled likelihood as an approximation. For the profiled likelihood we use the fact, that the derivative of the likelihood at its maximum is zero. Therefore one can construct from $\frac{\textrm{dL}}{\textrm{db}_i}=0$ the background rates $b_{i, max}$ that maximize the likelihood for a given model $m_{i}(\theta)$ and observed quantities. Solving $\frac{\textrm{dL}}{\textrm{db}_i}=0$ for b gives equation \ref{eq:bmax}.
\begin{equation}
	b_{i,max}(\theta)=\frac{1}{2(t_{b}+t_{d})}(B_{i}+D_{i}-m_{i}(\theta)(t_{b}+t_{d})+\sqrt{(B_{i}+D_{i}+m_{i}(\theta)(t_{b}+t_{d}))^{2}-4m_{i}(\theta)D_{i}(t_{d}-t_{b})}
  \label{eq:bmax}
\end{equation}

@@

\begin{figure}
    \begin{centering}
        \includegraphics{figures/electronic_noise.pdf}
        \caption{Count spectrum for detector 0 integrated over 1000 seconds. One can clearly see the feature in the Non-PSD events at 1.5 MeV that is not visible in the PSD events.}
        \label{fig:electronic}
    \end{centering}
\end{figure}

@@

\begin{figure}
  \begin{centering}
    \includegraphics[width=1\linewidth]{figures/spi_lightcurve_ts.pdf}
    \includegraphics[width=1\linewidth]{figures/gbm_lightcurve_ts.pdf}
    \caption{Light curve for GRB120711A.}
    \label{fig:time_selection}
  \end{centering}
\end{figure}

@@

\begin{appendix}
  \section{Model Checking Plots}
  \label{appendix1}
  In this section we show a selection of the PPC- and QQ-plots for the simultaneous synchrotron fit to the SPI and GBM data for GRB120711A.
  \begin{figure}[h]
    \begin{centering}
      \includegraphics[width=0.45\linewidth]{figures/n0_ppc.pdf}
      \includegraphics[width=0.45\linewidth]{figures/n0_qq.pdf}
      \caption{PPC (left) and QQ (right) plot for the GBM detector n0.}
      \label{fig:n0_model_check}
    \end{centering}
  \end{figure}
  \begin{figure}[h]
    \begin{centering}
      \includegraphics[width=0.45\linewidth]{figures/b0_ppc.pdf}
      \includegraphics[width=0.45\linewidth]{figures/b0_qq.pdf}
      \caption{PPC (left) and QQ (right) plot for the GBM detector b0.}
      \label{fig:b0_model_check}
    \end{centering}
  \end{figure}
  \begin{figure}[h]
    \begin{centering}
      \includegraphics[width=0.45\linewidth]{figures/low_13_ppc.pdf}
      \includegraphics[width=0.45\linewidth]{figures/low_13_qq.pdf}
      \caption{PPC (left) and QQ (right) plot for the low energy range of SPI detector 13 (all single events).}
      \label{fig:low_13_model_check}
    \end{centering}
  \end{figure}
  \begin{figure}[h]
    \begin{centering}
      \includegraphics[width=0.45\linewidth]{figures/psd_13_ppc.pdf}
      \includegraphics[width=0.45\linewidth]{figures/psd_13_qq.pdf}
      \caption{PPC (left) and QQ (right) plot for the middle energy range of SPI detector 13 (only PSD events).}
      \label{fig:psd_13_model_check}
    \end{centering}
  \end{figure}
\end{appendix}

\bibliographystyle{aa}
\bibliography{biblio}

>>>>

\begin{figure}
    \begin{centering}
        \includegraphics{figures/lightcurve.pdf}
        \caption{Lightcurve for SPI detector 0 for GRB120711A. The transient source is clearly visible as well as the constant background for the time when the transient is not active. The time intervals we use for the independent background observation is marked with the green shaded area.}
        \label{fig:lightcurve}
    \end{centering}
\end{figure}

@@

\begin{multline}
	L(\theta, b_{i}|D_{i}, B_{i},t_{d},t_{b}) = \frac{(t_{d}(m_{i}(\theta)+ b_{i}))^{D_{i}}}{D_{i}!}\cdot \\
  e^{-t_{d}(m_{i}(\theta)+b_{i})}\frac{(t_{b} b_{i})^{B_{i}}}{B_{i}!} e^{-t_{b} b_{i}}
  \label{eq:likelihood_m_and_b}
\end{multline}

Now we can marginalize over the parameters $b_{i}$, that we are not interested in and that we have no further information about a priori, despite the fact that they can not be negative, which leaves us with equation \ref{eq:likelihood_marg}.

\begin{multline}
	L(\theta|D_{i}, B_{i},t_{d},t_{b}) = \int_{0}^{\infty}\textrm{db}_{i}\frac{(t_{d}(m_{i}(\theta)+ b_{i}))^{D_{i}}}{D_{i}!}\cdot\\
  e^{-t_{d}(m_{i}(\theta)+b_{i})} \frac{(t_{b} b_{i})^{B_{i}}}{B_{i}!} e^{-t_{b}b_{i}}
  \label{eq:likelihood_marg}
\end{multline}

Because there is no easy analytical solution for this integral, one often uses a so-called profiled likelihood as an approximation. For the profiled likelihood we use the fact, that the derivative of the likelihood at its maximum is zero. Therefore one can construct from $\frac{\textrm{dL}}{\textrm{db}_i}=0$ the background rates $b_{i, max}$ that maximize the likelihood for a given model $m_{i}(\theta)$ and observed quantities. Solving $\frac{\textrm{dL}}{\textrm{db}_i}=0$ for b gives equation \ref{eq:bmax}.
\begin{multline}
	b_{i,max}(\theta)=\frac{1}{2(t_{b}+t_{d})}(B_{i}+D_{i}-m_{i}(\theta)(t_{b}+t_{d})+\\
  \sqrt{(B_{i}+D_{i}+m_{i}(\theta)(t_{b}+t_{d}))^{2}-4m_{i}(\theta)D_{i}(t_{d}-t_{b})}
  \label{eq:bmax}
\end{multline}

@@

\begin{figure}
    \begin{centering}
        \includegraphics{figures/electronic_noise.pdf}
        \caption{Count spectrum for detector 0 integrated over 1000 seconds. The Non-PSD event count spectrum is drawn in purple and the PSD events in grey. One can clearly see the feature in the Non-PSD events at 1.5 MeV (green shaded area) that is not visible in the PSD events.}
        \label{fig:electronic}
    \end{centering}
\end{figure}

@@

\begin{figure}
  \begin{centering}
    \includegraphics[width=1\linewidth]{figures/spi_lightcurve_ts.pdf}
    \includegraphics[width=1\linewidth]{figures/gbm_lightcurve_ts.pdf}
    \caption{Light curve for GRB120711A in one GBM and one SPI detector. The green shaded area marks the time intervals used for the independent background observation and the blue shaded area is the active time interval used in the fit.}
    \label{fig:time_selection}
  \end{centering}
\end{figure}

@@

\begin{appendix}
  \section{Model Checking Plots}
  \label{appendix1}
  In this section we show a selection of the PPC- and QQ-plots for the simultaneous synchrotron fit to the SPI and GBM data for GRB120711A.
  \begin{figure}[h]
    \begin{centering}
      \includegraphics[width=0.45\linewidth]{figures/n0_ppc.pdf}
      \includegraphics[width=0.45\linewidth]{figures/n0_qq.pdf}
      \caption{PPC (left) and QQ (right) plot for the GBM detector n0. The green and purple shaded area are the one and two sigma contours. In the PPC plots the grey curve is the detected count spectrum and in the QQ plots the expected 1:1 relation between the cumulative model and observed counts.}
      \label{fig:n0_model_check}
    \end{centering}
  \end{figure}
  \begin{figure}[h]
    \begin{centering}
      \includegraphics[width=0.45\linewidth]{figures/b0_ppc.pdf}
      \includegraphics[width=0.45\linewidth]{figures/b0_qq.pdf}
      \caption{PPC (left) and QQ (right) plot for the GBM detector b0. The green and purple shaded area are the one and two sigma contours. In the PPC plots the grey curve is the detected count spectrum and in the QQ plots the expected 1:1 relation between the cumulative model and observed counts.}
      \label{fig:b0_model_check}
    \end{centering}
  \end{figure}
  \begin{figure}[h]
    \begin{centering}
      \includegraphics[width=0.45\linewidth]{figures/low_13_ppc.pdf}
      \includegraphics[width=0.45\linewidth]{figures/low_13_qq.pdf}
      \caption{PPC (left) and QQ (right) plot for the low energy range of SPI detector 13 (all single events). The green and purple shaded area are the one and two sigma contours. In the PPC plots the grey curve is the detected count spectrum and in the QQ plots the expected 1:1 relation between the cumulative model and observed counts.}
      \label{fig:low_13_model_check}
    \end{centering}
  \end{figure}
  \begin{figure}[h]9
    \begin{centering}
      \includegraphics[width=0.45\linewidth]{figures/psd_13_ppc.pdf}
      \includegraphics[width=0.45\linewidth]{figures/psd_13_qq.pdf}
      \caption{PPC (left) and QQ (right) plot for the middle energy range of SPI detector 13 (only PSD events). The green and purple shaded area are the one and two sigma contours. In the PPC plots the grey curve is the detected count spectrum and in the QQ plots the expected 1:1 relation between the cumulative model and observed counts.}
      \label{fig:psd_13_model_check}
    \end{centering}
  \end{figure}
\end{appendix}

\bibliographystyle{aa}
\bibliography{bib}
